ABC - HTW Fits

### 2023-11-29 Model/abc_fit.qmd
abc_full_combo_group

abc_loss_plot <- function(abc_result,prior_samples)
{

  pgrid <- cbind(prior_samples,dist=abc_result$dist) |> mutate(dist=sqrt(dist))
  upper_limit <- quantile(pgrid$dist, 0.98)  # 95th percentile
  ggplot(pgrid, aes(x = c, y = lr, color = dist)) +
  geom_point() + 
 #scale_color_brewer(palette = "Spectral",limits = c(NA, upper_limit)) +
  #scale_color_viridis(limits = c(NA, upper_limit)) +
  scale_color_gradient(low = "blue", high = "red",limits = c(NA, upper_limit)) +  
  geom_contour(aes(x = c, y = lr, z = dist), color = 'white') +
  labs(title = "Point Plot of pgrid", x = "c", y = "lr") +
  theme_minimal()+ ggtitle("Prior Loss")

  post_grid <- cbind(prior_samples[abc_result$region,],dist=abc_result$dist[abc_result$region],k=abc_result$unadj.values)
  ggplot(post_grid, aes(x = c, y = lr, color = dist)) +
  geom_point() + 
  scale_color_gradient(low = "blue", high = "red",limits = c(NA, upper_limit)) +  
  geom_contour(aes(x = c, y = lr, z = dist), color = 'white') +
  labs(title = "Point Plot of pgrid", x = "c", y = "lr") +
  theme_minimal() + ggtitle("Posterior Loss")

@@

  # Generate posterior density plots
  postV <- ggplot(posterior_samples_long, aes(x=value)) +
    geom_density() +
    facet_wrap(~name, scales="free") +
    theme_minimal() +
    labs(x="Value", y="Density", title="Posterior Density Plots")
  print(postV)

@@

@@

  # Generate plot for simulation results
  plot_fsv <- fsv |> 
    tidyr::pivot_longer(y:pred, names_to="Resp", values_to="vx") |> 
    ggplot(aes(x, vx, fill=Resp, group=Resp)) +
    stat_summary(geom="bar", fun="mean", position=position_dodge()) +
    scale_fill_manual(values=col_themes$wes2) +
    scale_x_continuous(breaks=sort(unique(data$x)), labels=sort(unique(fsv$x))) +
    theme(legend.title = element_blank(), legend.position="top") +
    ggtitle("Fit to Test Only")
  print(plot_fsv)
}

@@

@@

dst_v_tetr <- avg_dsv[expMode2=="Test" | expMode2=="Train",]$y

abc_v_tetr <- abc(
  target = dst_v_tetr,
  param = prior_samples,
  sumstat =do.call(rbind, sd_v_tetr),
  tol = .01,
  method = "rejection",
  names=colnames(dst_v_tetr)
)

dst_v_te <- avg_dsv[expMode2=="Test",]$y
sd_v_te <- sd_v_tetr[85:90,]

abc_v_te <- abc(
  target = dst_v_te,
  param = prior_samples,
  sumstat =do.call(rbind, sd_v_te),
  tol = .01,
  method = "rejection",
  names=colnames(dst_v_te)
)

abc_plot(abc_v_tetr, avg_dsv) + abc_loss_plot(abc_v_tetr,prior_samples)
abc_plot(abc_v_te, avg_dsv) + abc_loss_plot(abc_v_te,prior_samples)

pgrid |> filter(dist<80) |> ggplot(aes(x=c,y=lr,col=dist)) + geom_point() + viridis::scale_color_viridis()

pgrid |> filter(dist<1700) |> ggplot(aes(x=c,y=dist)) + geom_line()
pgrid |> filter(dist<1700) |> ggplot(aes(x=c,y=dist)) + geom_smooth()

str(pgrid)
ggplot(pgrid)+
  geom_tile(aes(x = c, y = lr, fill = dist)) +
  geom_contour(aes(x = alpha, y = beta, z = nLL), color = 'white')

pgrid |> filter(dist<50) |> ggplot(aes(x = c, y = lr, col = factor(dist))) +
  geom_point() +
  viridis::scale_color_viridis(discrete = TRUE)

upper_limit <- quantile(pgrid$dist, 0.98)  # 95th percentile
ggplot(pgrid, aes(x = c, y = lr, color = dist)) +
  geom_point() +
 #scale_color_brewer(palette = "Spectral",limits = c(NA, upper_limit)) +
  #scale_color_viridis(limits = c(NA, upper_limit)) +
  scale_color_gradient(low = "blue", high = "red",limits = c(NA, upper_limit)) +  
  geom_contour(aes(x = c, y = lr, z = dist), color = 'white') +
  labs(title = "Point Plot of pgrid", x = "c", y = "lr") +
  theme_minimal()

ggplot(pgrid |>filter(dist<66)) +
    geom_tile(aes(x = c, y = lr, fill = dist)) +
    geom_contour(aes(x = c, y = lr, z = dist)) +
    scale_fill_viridis() +
    labs(title = "Your Plot Title", x = "c", y = "lr") +
    theme_minimal()

pgrid |> ggplot(aes(x=c,y=lr,fill=as.factor(dist))) + geom_point() + viridis::scale_fill_viridis()

pgrid |> ggplot(aes(x = c, y = lr, col = factor(dist))) +
  geom_point() +
  viridis::scale_color_viridis(discrete = TRUE)

# 
# dim(simulated_data$y)[2]
# dim(sim)
# dim(t(sim))
```

@@

@@

abc_plot(abc_c_tetr, avg_dsc)
abc_plot(abc_c_te, avg_dsc)

>>>>

abc_loss_plot <- function(abc_result,prior_samples)
{
  pgrid <- cbind(prior_samples,dist=abc_result$dist) |> mutate(dist=sqrt(dist))
  upper_limit <- quantile(pgrid$dist, 0.95)  # 95th percentile
  pgrid <- pgrid |> filter(dist <= upper_limit)
  gg_prior <- ggplot(pgrid, aes(x = c, y = lr, color = dist)) +
  geom_point() + 
  #scale_color_viridis(limits = c(NA, upper_limit)) +
  scale_color_gradient(low = "blue", high = "red",limits = c(NA, upper_limit)) +  
  geom_contour(aes(x = c, y = lr, z = dist), color = 'white') +
  labs(title = "Point Plot of pgrid", x = "c", y = "lr") +
  theme_minimal()+ ggtitle("Prior Loss")

  post_grid <- cbind(prior_samples[abc_result$region,],dist=abc_result$dist[abc_result$region],k=abc_result$unadj.values)
  gg_post <- ggplot(post_grid, aes(x = c, y = lr, color = dist)) +
  geom_point() + 
  scale_color_gradient(low = "blue", high = "red",limits = c(NA, upper_limit)) +  
  geom_contour(aes(x = c, y = lr, z = dist), color = 'white') +
  labs(title = "Point Plot of pgrid", x = "c", y = "lr") +
  theme_minimal() + ggtitle("Posterior Loss")
  gg_prior+gg_post

@@

  # Generate posterior density plots
  postV <- ggplot(posterior_samples_long, aes(x=value)) +
    geom_density() +
    facet_wrap(~name, scales="free") +
    theme_minimal() +
    labs(x="Value", y="Density", title="Posterior Density Plots")

@@

  print(paste0("Mode of c: ", mode_c, ", Mode of lr: ", mode_lr))

@@

  # Generate plot for simulation results
  plot_fsv <- fsv |> 
    tidyr::pivot_longer(y:pred, names_to="Resp", values_to="vx") |> 
    ggplot(aes(x, vx, fill=Resp, group=Resp)) +
    stat_summary(geom="bar", fun="mean", position=position_dodge()) +
    scale_fill_manual(values=col_themes$wes2) +
    scale_x_continuous(breaks=sort(unique(data$x)), labels=sort(unique(fsv$x))) +
    theme(legend.title = element_blank(), legend.position="top") +
    ggtitle("Fit to Test Only")

  postV + plot_fsv
}

@@

# General function for ABC fits with model flexibility
run_abc_fits <- function(data, input_layer, output_layer, simulation_function, n_prior_samples = 1000, tol = 0.01, return_dat = "train_data,test_data") {

  # Generate prior samples
  prior_samples <- generate_prior_c_lr(n_prior_samples)

    # Convert the string to a function
  if (is.character(simulation_function)) {
    simulation_function <- get(simulation_function, mode = "function")
  } else {
    simulation_function <- match.fun(simulation_function)
  }

  plan(multisession)
  # Run simulations for each prior sample
  simulation_results <- future_map_dfc(seq_len(nrow(prior_samples)), function(idx) {
    params <- prior_samples[idx, ]
    simulation_function(data=as.data.table(data), c=params$c, lr=params$lr, input_layer=input_layer, output_layer=output_layer, return_dat = return_dat)
  }, .options = furrr_options(seed = TRUE))

  # Extract target data
  target_data_train_test <- data[expMode2 %in% c("Test", "Train"), ]$y
  target_data_test <- data[expMode2 == "Test", ]$y

  # ABC for train and test data
  abc_train_test <- abc(
    target = target_data_train_test,
    param = prior_samples,
    sumstat = do.call(rbind, simulation_results),
    tol = tol,
    method = "rejection"
  )

  # ABC for test data only
  abc_test <- abc(
    target = target_data_test,
    param = prior_samples,
    sumstat = do.call(rbind, simulation_results[85:90, ]),
    tol = tol,
    method = "rejection"
  )

  # Return results
  list(abc_train_test = abc_train_test, abc_test = abc_test)
}

# Example usage for different models
n_priors=20000
abc_results_varied_exam <- run_abc_fits(avg_dsv, input_layer, output_layer, full_sim_exam, n_prior_samples = n_priors)
abc_results_varied_alt_exam <- run_abc_fits(avg_dsv, input_layer, output_layer, full_sim_alt_exam, n_prior_samples = n_priors)
abc_results_varied_alm <- run_abc_fits(avg_dsv, input_layer, output_layer, full_sim_alm, n_prior_samples = n_priors)

abc_results_constant_alm <- run_abc_fits(avg_dsc, input_layer, output_layer, full_sim_alm, n_prior_samples = n_priors)
abc_results_constant_exam <- run_abc_fits(avg_dsc, input_layer, output_layer, full_sim_exam, n_prior_samples = n_priors)
abc_results_constant_alt_exam <- run_abc_fits(avg_dsc, input_layer, output_layer, full_sim_alt_exam, n_prior_samples = n_priors)

# store all models in list and save to data/model_cache
model_cache <- tibble:lst(
  abc_results_varied_exam = abc_results_varied_exam,
  abc_results_varied_alt_exam = abc_results_varied_alt_exam,
  abc_results_varied_alm = abc_results_varied_alm,
  abc_results_constant_alm = abc_results_constant_alm,
  abc_results_constant_exam = abc_results_constant_exam,
  abc_results_constant_alt_exam = abc_results_constant_alt_exam
)

saveRDS(model_cache,here::here(paste0("data/model_cache/abc_group_",format(Sys.time(), "%H_%M_%OS"),".rds")))

n_priors=20000
args_list <- tibble::lst(
  abc_v_exam=list(data = avg_dsv, input_layer = input_layer, output_layer = output_layer, simulation_function = full_sim_exam, n_prior_samples = n_priors),
  abc_v_alt_exam=list(data = avg_dsv, input_layer = input_layer, output_layer = output_layer, simulation_function = full_sim_alt_exam, n_prior_samples = n_priors),
  abc_v_alm=list(data = avg_dsv, input_layer = input_layer, output_layer = output_layer, simulation_function = full_sim_alm, n_prior_samples = n_priors),
  abc_c_alm=list(data = avg_dsc, input_layer = input_layer, output_layer = output_layer, simulation_function = full_sim_alm, n_prior_samples = n_priors),
  abc_c_exam=list(data = avg_dsc, input_layer = input_layer, output_layer = output_layer, simulation_function = full_sim_exam, n_prior_samples = n_priors),
  abc_c_alt_exam=list(data = avg_dsc, input_layer = input_layer, output_layer = output_layer, simulation_function = full_sim_alt_exam, n_prior_samples = n_priors)
)

results <- future_map(args_list, ~do.call(run_abc_fits, .x))
saveRDS(results,here::here(paste0("data/model_cache/abc_group12_",format(Sys.time(), "%H_%M_%OS"),".rds")))

@@

dst_v_tetr <- avg_dsv[expMode2=="Test" | expMode2=="Train",]$y
abc_v_tetr <- abc(
  target = dst_v_tetr,
  param = prior_samples,
  sumstat =do.call(rbind, sd_v_tetr),
  tol = .01,
  method = "rejection",
  names=colnames(dst_v_tetr)
)

dst_v_te <- avg_dsv[expMode2=="Test",]$y
sd_v_te <- sd_v_tetr[85:90,]
abc_v_te <- abc(
  target = dst_v_te,
  param = prior_samples,
  sumstat =do.call(rbind, sd_v_te),
  tol = .01,
  method = "rejection",
  names=colnames(dst_v_te)
)

abc_plot(abc_v_tetr, avg_dsv) / { abc_loss_plot(abc_v_tetr,prior_samples) }
abc_plot(abc_v_te, avg_dsv) / { abc_loss_plot(abc_v_te,prior_samples) }

@@

abc_plot(abc_c_tetr, avg_dsc)

@@

abc_plot(abc_c_tetr, avg_dsc) / { abc_loss_plot(abc_c_tetr,prior_samples) }
abc_plot(abc_c_te, avg_dsc) / { abc_loss_plot(abc_c_te,prior_samples) }
